---
title: "R Notebook"
output: html_notebook
---

### PURPOSE

To demonstrate how to work with arrayed UK Biobank files

### Setup
```{r}


#install.packages("tidyverse")
#install.packages("data.table")

library(tidyverse)
library(data.table)


# How to show the documentation for the function
?fread

```
#---------------------------------

### BEGIN

### 1. Start with dates field
```{r}

# This is a good way to have a complete participant list

all_f.eids <- fread("~/Documents/UKB Data/f.53.tab")  %>%                       
                  select(f.eid)

dim(all_f.eids)
# 502411      1
```

### 2. Read the coding table
```{r}

## ICD 10

# The coding reference table
coding19<-     fread("~/Documents/UKB Data/coding19.tsv") %>% 
                   select(coding, meaning)
```


### 3. Read data inputs
```{r}

# File containing codes

code_records<- fread("~/Documents/UKB Data/f.41270.tab") 

# Show the size of the object (rows, cols)
dim(code_records)
# 502411    244
 

# File containing dates

date_records<- fread("~/Documents/UKB Data/f.41280.tab") 

dim(date_records)
# 502411    244


# These two fields are companions, they are intended to join by f.eid and array index
```

### 4. Reshape coding file from wide file to long format
```{r}
## Convert both code records and date records to long format
## we have multiple records per person.

codes_long<- code_records %>%                    
               melt(id.vars = "f.eid") %>%     # Convert wide format to long format (retain f.eid)
              .[value != ""]                   # Filter out empty space (this is a data.table filter)

dim(codes_long)

# 6302095       3
```

I'm really sorry for mixing data.table filters with tidyverse syntax, its just that data.table functions work a lot better with large objects

### 5. Filter only to our desired codes
```{r}
mini<- codes_long[value %like% "C50|D05"] 

dim(mini)

# 25712     3

# How many unique people?
length(unique(mini$f.eid)) 

# 19864
```


```{r}
## Extract index array (needed for joining)

mini<- mini %>% mutate(array_idx= str_sub(variable, 8)) %>%
                rename(coding= value) %>%
                select(-variable)
```

### 6. Optional enrich with meaning text
```{r}
## Join with coding table (enrich with meanings)

mini<- left_join(mini, coding19)

head(mini)

#      f.eid coding array_idx                              meaning
# 1: 1000874   C509      .0.0            C50.9 Breast, unspecified
# 2: 1001025   C509      .0.0            C50.9 Breast, unspecified
# 3: 1001241   C509      .0.0            C50.9 Breast, unspecified
# 4: 1001262   C504      .0.0 C50.4 Upper-outer quadrant of breast
# 5: 1001359   C509      .0.0            C50.9 Breast, unspecified
# 6: 1002125   C504      .0.0 C50.4 Upper-outer quadrant of breast
```


### 7. Reshape dates (wide to long)

```{r}
dates_long<-  date_records %>% 
                   melt(id.vars = "f.eid") %>%                  # Convert wide format to long format
                   .[!is.na(value)] %>%                         # filter out empty space
                   mutate(array_idx= str_sub(variable, 8),      # extract array index for joining
                               date= ymd(value)) %>%            # convert text to date
                   select(-variable)

head(dates_long)
#      f.eid      value array_idx       date
# 1: 1000874 1997-04-15      .0.0 1997-04-15
# 2: 1000923 2010-06-28      .0.0 2010-06-28
# 3: 1001025 1998-06-28      .0.0 1998-06-28
# 4: 1001241 2013-10-15      .0.0 2013-10-15
# 5: 1001262 1999-08-26      .0.0 1999-08-26
# 6: 1001359 2000-10-30      .0.0 2000-10-30
```


### 8. Combine codes and dates
```{r}
# Combine codes with dates

combined<- inner_join(mini, dates_long); 
dim(combined)
# 25712     6

length(unique(combined$f.eid))
# 19864
```
### 9. Find the minimum date
```{r}

# Find first occurrence of either C50 or D05
# And to create One row per person

t0<- combined %>% 
  group_by(f.eid) %>% 
  summarise(minDate= min(date)) %>%
  mutate(status= 1)

head(t0)


#     f.eid minDate    status
#     <int> <date>      <dbl>
# 1 1000874 1997-04-15      1
# 2 1000923 2001-11-22      1
# 3 1001025 1998-06-28      1
# 4 1001241 2013-10-15      1
# 5 1001262 1999-08-26      1
# 6 1001359 2000-10-30      1
```
### 10. Create output object
```{r}
out<- full_join(all_f.eids, t0) %>%
      set_names(c("f.eid", "Breast_date", "Breast_status")) %>%
      mutate(Breast_status= ifelse(is.na(Breast_status), 0, Breast_status))

head(out)
#      f.eid Breast_date Breast_status
# 1: 6026576        <NA>             0
# 2: 6026582        <NA>             0
# 3: 1000018        <NA>             0
# 4: 1000020        <NA>             0
# 5: 1000034        <NA>             0
# 6: 1000041        <NA>             0


sum(out$Breast_status)
 
# 19864
```


#=====================
### SAVE
```{r}
saveRDS(out, "Breast_status.rds")

write_tsv(out, "Breast_status.tsv")
```

#-----------------------

# END
