---
title: "Descriptive Statistics and Data Analysis"
author: "Gracia Andriamiadana"
date: "02/08/2023"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(dplyr)
library(magrittr)
library(ggplot2)
library(lubridate)
library(broom.mixed)
library(vtable)
library(survival)
library(broom)
library(purrr)
library(arsenal)

setwd("/Users/BHCadmin/Documents/mres_gracia/")
```

## Load and edit data tables
```{r}
demographic_table <- read.csv("upd_demographics_table_UKB_data.tsv", header = T, sep = "\t")
cmr_measures_tbl <- read.csv("CMR_patients_in_UKB.tsv", header = T, sep = "\t")
cmr_original_df <- read.csv('cmr_measures_updated.csv', header = T, sep = ';') 
cmr_t1_df <- read.csv('t1_qcd_evan.csv', header = T, sep = ',') %>% rename(f.eid = feid)
bsa_tbl <- read.csv("BSA_of_UKB_patients.tsv", header = T, sep = "\t")
bmi_tbl <- read.csv("BMI_of_patients_in_UKB.tsv", sep = "\t")
diabetes_tbl = read.csv("updated_diabetes_tbl.tsv", sep = "\t")
hypertension_tbl = read.csv("updated_hypertension_tbl.tsv", sep = "\t")
cholesterol_tbl = read.csv("updated_cholesterol_tbl.tsv", sep = "\t")

IHD_tbl = read.csv("updated_IHD_tbl.tsv", sep = "\t")
non_IHD_tbl = read.csv("updated_non_IHD_tbl.tsv", sep = "\t")
HF_tbl = read.csv("updated_HF_tbl.tsv", sep = "\t")
HA_tbl = read.csv("updated_HA_tbl.tsv", sep = "\t")
pericardial_tbl <- read.csv("updated_pericardial_tbl.tsv", sep = "\t")
stroke_tbl = read.csv("updated_stroke_tbl.tsv", sep = "\t")

BC_tbl <- read.csv("updated_BC_tbl.tsv", sep = "\t", header = T)
BC_ids <- BC_tbl %>% dplyr::filter(prevalent_status == 1) %>% select(f.eid)
nrow(BC_ids)
BC_ids <- as.vector(BC_ids$f.eid)

non_BC_ids <- BC_tbl %>% dplyr::filter(BC_status == 0) %>% select(f.eid)
nrow(non_BC_ids)
non_BC_ids <- as.vector(non_BC_ids$f.eid)

RT_tbl <- read.csv("updated_RT_tbl.tsv", sep = "\t", header = T)
RT_ids <- RT_tbl %>% dplyr::filter(RT_status == 1) %>% select(f.eid)
nrow(RT_ids)

RT_and_BC_ids <- intersect(as.vector(BC_ids), as.vector(RT_ids$f.eid))
length(RT_and_BC_ids)

no_RT_ids <- RT_tbl %>% dplyr::filter(RT_status == 0) %>% select(f.eid)
no_RT_and_BC_ids <- intersect(as.vector(BC_ids), as.vector(no_RT_ids$f.eid))
length(no_RT_and_BC_ids)

cmr_ids <- cmr_measures_tbl$f.eid
length(cmr_ids)
cmr_no_bc_ids <- intersect(cmr_ids, non_BC_ids)
length(cmr_no_bc_ids)

cmr_bc_ids <- intersect(cmr_ids, BC_ids)
length(cmr_bc_ids)

cmr_bc_rt_ids <- intersect(cmr_ids, RT_and_BC_ids)
length(cmr_bc_rt_ids)

cmr_bc_but_no_mrt <- intersect(cmr_ids, no_RT_and_BC_ids)
length(cmr_bc_but_no_mrt)

sample_df <- read.csv("analysis_sample_UKB_patients.tsv", sep = "\t")

# GLM_df <- readRDS("table_for_regression_models.rds")

```

## Data Preprocessing and Wrangling for summary stats

```{r preprocessing_of_demographics_table}
tmp <- demographic_table %>% dplyr:: select(contains("status"))
factor_cols <- names(tmp)

tmp <- demographic_table %>% dplyr:: select(contains("alc"))
factor_cols <- c(factor_cols, names(tmp))

df1 <- demographic_table %>% dplyr:: select(contains("ethnicity"))
factor_cols <- c(factor_cols, names(df1))

demographic_table <- demographic_table %>% mutate(across(all_of(factor_cols), factor))
demographic_table$recruitment_date <- ymd(demographic_table$recruitment_date)
demographic_table$sex <- factor(demographic_table$sex)
demographic_table$f.eid <- as.character(demographic_table$f.eid)

demographic_table$ethnicity_at_recruitment[demographic_table["ethnicity_at_recruitment"] != 1] <- 0
# demographic_table$ethnicity_at_recruitment[demographic_table["ethnicity_at_recruitment"] == -1000] <- NA
demographic_table$ethnicity_at_calib_visit[demographic_table["ethnicity_at_calib_visit"] == -1000] <- NA
demographic_table$ethnicity_at_imaging_visit1[demographic_table["ethnicity_at_imaging_visit1"] == -1000] <- NA

stroke_tbl_BC <- stroke_tbl %>% # dplyr::filter(f.eid %in% BC_ids) %>% 
  dplyr::select(f.eid, prevalent_status, incident_status) %>% rename(prev_status_stroke = prevalent_status,
                                                                     incident_status_stroke = incident_status)

IHD_tbl_BC <- IHD_tbl %>% # dplyr::filter(f.eid %in% BC_ids) %>% 
  dplyr::select(f.eid, prevalent_status, incident_status) %>% rename(prev_status_IHD = prevalent_status,
                                                                     incident_status_IHD = incident_status)

NICM_tbl_BC <- non_IHD_tbl %>% # dplyr::filter(f.eid %in% BC_ids) %>% 
  dplyr::select(f.eid, prevalent_status, incident_status) %>% rename(prev_status_NICM = prevalent_status,
                                                                     incident_status_NICM = incident_status)

pericard_tbl_BC <- pericardial_tbl %>% # dplyr::filter(f.eid %in% BC_ids) %>% 
  dplyr::select(f.eid, prevalent_status, incident_status) %>% rename(prev_status_pericard = prevalent_status,
                                                                     incident_status_pericard = incident_status)

HF_tbl_BC <- HF_tbl %>% # dplyr::filter(f.eid %in% BC_ids) %>% 
  dplyr::select(f.eid, prevalent_status, incident_status) %>% rename(prev_status_HF = prevalent_status,
                                                                     incident_status_HF = incident_status)

myocard_tbl_BC <- HA_tbl %>% # dplyr::filter(f.eid %in% BC_ids) %>% 
  dplyr::select(f.eid, prevalent_status, incident_status) %>% rename(prev_status_myocard_inf = prevalent_status,
                                                                     incident_status_myocard_inf = incident_status)

diabetes_tbl_BC <- diabetes_tbl %>% # dplyr::filter(f.eid %in% BC_ids) %>% 
  dplyr::select(f.eid, prevalent_status, incident_status) %>% 
  rename(prev_status_diabetes = prevalent_status, 
         incident_status_diabetes = incident_status)
 

cholesterol_tbl_BC <- cholesterol_tbl %>% # dplyr::filter(f.eid %in% BC_ids) %>% 
  dplyr::select(f.eid, prevalent_status, incident_status) %>% 
  rename(prev_status_cholesterol = prevalent_status, 
         incident_status_cholesterol = incident_status)

hypertension_tbl_BC <- hypertension_tbl %>% # dplyr::filter(f.eid %in% BC_ids) %>% 
  dplyr::select(f.eid, prevalent_status, incident_status) %>% 
  rename(prev_status_hypertension = prevalent_status, 
         incident_status_hypertension = incident_status)

demographic_table$f.eid <- as.integer(demographic_table$f.eid)

df_list <- list(BC_tbl %>% select(-recruitment_date), 
                RT_tbl %>% select(-recruitment_date, -prevalent_status, -incident_status, -RT_session_date),
                demographic_table, bmi_tbl, diabetes_tbl_BC, hypertension_tbl_BC, cholesterol_tbl_BC,
                IHD_tbl_BC, NICM_tbl_BC, HF_tbl_BC, stroke_tbl_BC, pericard_tbl_BC, myocard_tbl_BC
                )

GLM_df <- df_list %>% reduce(full_join, by='f.eid')

tmp <- GLM_df %>% dplyr:: select(contains("status"))
factor_cols <- names(tmp)

tmp <- GLM_df %>% dplyr:: select(contains("alc"))
factor_cols <- c(factor_cols, names(tmp))

df1 <- GLM_df %>% dplyr:: select(contains("ethnicity"))
factor_cols <- c(factor_cols, names(df1))

GLM_df <- GLM_df %>% mutate(across(all_of(factor_cols), factor))
# data<-replace(df$Marks, df$Marks<0, 0)
GLM_df$recruitment_date <- ymd(GLM_df$recruitment_date)
GLM_df$BC_diagnosis_date <- ymd(GLM_df$BC_diagnosis_date)
# GLM_df$RT_session_date <- ymd(GLM_df$RT_session_date)
GLM_df$sex <- factor(GLM_df$sex)
GLM_df$time = as.numeric(difftime(time2 = GLM_df$recruitment_date, time1 = GLM_df$BC_diagnosis_date))
GLM_df$time[is.na(GLM_df$time)] <- 0

GLM_df <- GLM_df %>%
  mutate(
    alc_grouped = case_when(
      alc_intake_at_recruitment %in% c("6") ~ "Never",     # Group A and B into "Group1"
      alc_intake_at_recruitment %in% c("4", "5") ~ "Less than once a week",     # Group C and D into "Group2"
      alc_intake_at_recruitment %in% c("1", "2", "3") ~ "More than once a week",     # Group C and D into "Group2"
      alc_intake_at_recruitment %in% c("-3") ~ NA,     # Group C and D into "Group2"
      TRUE ~ alc_intake_at_recruitment                           # Keep other categories as-is
    )
  ) %>%
  mutate(
    smoking_grouped = case_when(
      smoking_status_at_recruitment %in% c("0", "1") ~ "Not smoking",     # Group C and D into "Group2"
      smoking_status_at_recruitment %in% c("2") ~ "Currently smoking",     # Group C and D into "Group2"
      smoking_status_at_recruitment %in% c("-3") ~ NA,     # Group C and D into "Group2"
      TRUE ~ smoking_status_at_recruitment                           # Keep other categories as-is
    )
  )

# nrow(GLM_df)
head(GLM_df, 5)

# write.table(GLM_df, file = "table_for_regression_models.tsv", sep = "\t")
# saveRDS(GLM_df, file = "table_for_regression_models.rds")
# GLM_df <- readRDS("table_for_regression_models.rds")

# GLM_df$followup = time_length(difftime(time1 = ymd("2023-05-01"), time2 = GLM_df$recruitment_date), "years")
# GLM_df$followup2 = time_length(difftime(time1 = ymd("2023-05-31"), time2 = GLM_df$recruitment_date), "years")
# summary(GLM_df %>% select(followup, followup2))
```

## Baseline Characteristics of Whole and Sub Samples

### CVDs and Demographics of CMR Sub Samples

```{r cmr_descriptive_stats}

df_cmr_baseline <- inner_join(cmr_measures_tbl, GLM_df, by = "f.eid")

test_df <- df_cmr_baseline %>% dplyr::filter(f.eid %in% cmr_bc_ids) # CMR and BC 
test_df <- df_cmr_baseline %>% dplyr::filter(f.eid %in% cmr_no_bc_ids)
nrow(test_df)

table_for_summary_stats <- tableby(~BC_status + age_at_recruitment + sex + ethnicity_at_recruitment  + education_score + deprivation_index + alc_grouped + smoking_grouped  + mets_imaging_visit1 + prev_status_diabetes + incident_status_diabetes + prev_status_hypertension + incident_status_hypertension + prev_status_cholesterol + incident_status_cholesterol + prev_status_IHD + incident_status_IHD + prev_status_NICM + incident_status_NICM + prev_status_HF + incident_status_HF + prev_status_stroke + incident_status_stroke + prev_status_pericard + incident_status_pericard + prev_status_myocard_inf + incident_status_myocard_inf, data = test_df,
                      numeric.stats=c("median", "q1q3"), na.tableby(FALSE))

sum_table <- summary(table_for_summary_stats, text = NULL)
sum_table <- as.data.frame(sum_table)
View(sum_table)

# write.table(sum_table, 'summary_stats_CMR-BC_patients.tsv', sep = "\t")
write.table(sum_table, 'summary_stats_CMR-no-BC_patients.tsv', sep = "\t")
```

```{r}
cmr_df <- full_join(cmr_measures_tbl, cmr_original_df, by = "f.eid")
nrow(cmr_df)
cmr_df <- full_join(cmr_df, cmr_t1_df %>% select(f.eid, t1), by = "f.eid")
nrow(cmr_df)
cmr_df <- full_join(cmr_df, GLM_df, by = "f.eid")
nrow(cmr_df)
cmr_df <- full_join(cmr_df, bsa_tbl, by = "f.eid")
nrow(cmr_df)
cmr_df$t1[cmr_df$t1=="--"] <- NA
cmr_df$t1 <- as.numeric(cmr_df$t1)

cmr_df <- cmr_df %>% mutate(LVSVi = LVSV..mL./bsa_imaging_visit1,
                            max_LAVi = LAV.max..mL./bsa_imaging_visit1,
                            RVEDVi = RVEDV..mL./bsa_imaging_visit1,
                            RVSVi = RVSV..mL./bsa_imaging_visit1,
                            LAEF = LAEF....,
                            LVM_LDEDV = LVM..g./LVEDV..mL.,
                            RVEF = RVEF....,
                            DAo = DAo.distensibility..10.3.mmHg.1.
                            )

selected_cols <- cmr_df[, grepl("WT_AHA", names(cmr_df))]
result <- do.call(pmax, c(selected_cols, na.rm = TRUE))
cmr_df$max_LV_WT <- result

# colnames(cmr_df)
 
# test_df <- cmr_df %>% dplyr::filter(f.eid %in% cmr_bc_ids) 
test_df <- cmr_df %>% dplyr::filter(f.eid %in% cmr_no_bc_ids)
length(unique(test_df$f.eid))
nrow(test_df)
length(cmr_bc_ids)
length(cmr_no_bc_ids)

summary_cmr_measures <- tableby(~ max_LV_WT + LV_mass2 + LVEDVI + LVM_LDEDV + LVSVi + LVEF + LVGFI + GLS + t1 +
                                  max_LAVi + LAEF + RVEDVi + RVSVi + RVEF.... + DAo.distensibility..10.3.mmHg.1. ,
        numeric.stats=c( "median", "q1q3"), na.tableby(FALSE),
        data = test_df, )
summary_cmr_measures <- summary(summary_cmr_measures, text = NULL, digits=2) %>% as.data.frame()
View(summary_cmr_measures)
# write.table(summary_cmr_measures, 'summary_stats_CMR_measures-BCpatients.tsv', sep = "\t")
# write.table(summary_cmr_measures, 'summary_stats_CMR_measures-noBCpatients.tsv', sep = "\t")
```

## Logistic Regression Models for CVD outcomes (BC vs no BC)

```{r log_regressions}

# age, smoking, deprivation, alcohol, IPAQ activity level, diabetes, hypertension, high cholesterol
cox_models <- c()

### IHD
#### Model 2
test_df <- GLM_df %>% 
  select(f.eid, BC_status, BC_diagnosis_date, incident_status, prev_status_IHD, incident_status_IHD, 
         age_at_recruitment, sex, recruitment_date, time, bmi_recruitment,
         smoking_status_at_recruitment, ethnicity_at_recruitment,
         deprivation_index, alc_intake_at_recruitment, mets_imaging_visit1,
         prev_status_diabetes, prev_status_hypertension, prev_status_cholesterol) %>% filter(prev_status_IHD == 0 & 
                                                                                               incident_status == 0)

formula = (incident_status_IHD) ~ BC_status + 
  age_at_recruitment + 
  sex + ethnicity_at_recruitment + 
  smoking_status_at_recruitment + 
  alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = GLM_df, family = "binomial")
summary(model1)
tmp <- list(model1)
names(tmp) <- "IHD-Model2"
cox_models <- c(cox_models, tmp)

### Stroke
#### Model 2
test_df <- GLM_df %>% 
  select(f.eid, BC_status, BC_diagnosis_date, incident_status, prev_status_stroke, incident_status_stroke, 
         age_at_recruitment, sex, recruitment_date, time, bmi_recruitment,
         smoking_status_at_recruitment, ethnicity_at_recruitment,
         deprivation_index, alc_intake_at_recruitment, mets_imaging_visit1,
         prev_status_diabetes, prev_status_hypertension, prev_status_cholesterol) %>% filter(prev_status_stroke == 0 & 
                                                                                               incident_status == 0)

formula = incident_status_stroke ~ BC_status + 
  age_at_recruitment + 
  sex + ethnicity_at_recruitment + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = GLM_df, family = "binomial")
# summary(model1)
tmp <- list(model1)
names(tmp) <- "Stroke-Model2"
cox_models <- c(cox_models, tmp)

### HEART FAILURE
#### Model 2
test_df <- GLM_df %>% 
  select(f.eid, BC_status, BC_diagnosis_date, incident_status, prev_status_HF, incident_status_HF, 
         age_at_recruitment, sex, recruitment_date, time, bmi_recruitment,
         smoking_status_at_recruitment, ethnicity_at_recruitment,
         deprivation_index, alc_intake_at_recruitment, mets_imaging_visit1,
         prev_status_diabetes, prev_status_hypertension, prev_status_cholesterol) %>% filter(prev_status_HF == 0 & 
                                                                                               incident_status == 0)

formula = (incident_status_HF) ~ BC_status + 
  age_at_recruitment + 
  sex + ethnicity_at_recruitment + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = GLM_df, family = "binomial")
# summary(model1)
tmp <- list(model1)
names(tmp) <- "HF-Model2"
cox_models <- c(cox_models, tmp)

### NICM
#### Model 2
test_df <- GLM_df %>% 
  select(f.eid, BC_status, BC_diagnosis_date, incident_status, prev_status_NICM, incident_status_NICM, 
         age_at_recruitment, recruitment_date, time, sex, 
         smoking_status_at_recruitment, ethnicity_at_recruitment,
         deprivation_index, alc_intake_at_recruitment, mets_imaging_visit1,
         prev_status_diabetes, prev_status_hypertension, prev_status_cholesterol) %>% filter(prev_status_NICM == 0 & 
                                                                                               incident_status == 0)

formula = (incident_status_NICM) ~ BC_status + 
  age_at_recruitment + 
  sex + ethnicity_at_recruitment + 
  smoking_status_at_recruitment + 
  alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index  +
  prev_status_diabetes + 
  prev_status_cholesterol + 
  prev_status_hypertension
model1 <- glm(formula, data = GLM_df, family = "binomial")
summary(model1)
tmp <- list(model1)
names(tmp) <- "NICM-Model2"
cox_models <- c(cox_models, tmp)

### PERICARDIAL DISEASE
#### Model 2
test_df <- GLM_df %>% 
  select(f.eid, BC_status, BC_diagnosis_date, incident_status, prev_status_pericard, incident_status_pericard, 
         age_at_recruitment, sex, recruitment_date, time, bmi_recruitment,
         smoking_status_at_recruitment, ethnicity_at_recruitment,
         deprivation_index, alc_intake_at_recruitment, mets_imaging_visit1,
         prev_status_diabetes, prev_status_hypertension, prev_status_cholesterol) %>% filter(prev_status_pericard == 0& 
                                                                                               incident_status == 0)

formula = (incident_status_pericard) ~ BC_status + 
  age_at_recruitment + 
  sex + ethnicity_at_recruitment + 
  smoking_status_at_recruitment + 
  alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = GLM_df, family = "binomial")
# summary(model1)
tmp <- list(model1)
names(tmp) <- "Pericard-Model2"
cox_models <- c(cox_models, tmp)


### MYOCARDIAL INFARCTION
#### Model 2
test_df <- GLM_df %>% 
  select(f.eid, BC_status, BC_diagnosis_date, incident_status, prev_status_myocard_inf, incident_status_myocard_inf, 
         age_at_recruitment, sex, recruitment_date, time, bmi_recruitment,
         smoking_status_at_recruitment, ethnicity_at_recruitment,
         deprivation_index, alc_intake_at_recruitment, mets_imaging_visit1,
         prev_status_diabetes, prev_status_hypertension, prev_status_cholesterol) %>% filter(prev_status_myocard_inf == 0 & 
                                                                                               incident_status == 0)
# model1 <- glm(formula, data = cmr_df, family = "binomial")
# logistic regression instead of cox regression for now
# because time to events are missing [myocardial infarction dates etc.]

formula = incident_status_myocard_inf ~ BC_status + 
  age_at_recruitment + 
  sex + ethnicity_at_recruitment + 
  smoking_status_at_recruitment + 
  alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = GLM_df, family = "binomial")
summary(model1)
tmp <- list(model1)
names(tmp) <- "MYOINF-Model2"
cox_models <- c(cox_models, tmp)
```

#### Store log. regression results (BC vs no BC)

```{r}

# Step 1: Create a list of Cox models (you should have your own list)
print(length(cox_models))

# Specify the list of terms (variables) you want to include in the final merged table
# selected_terms <- c("BC_status1", "sex", "age_at_recruitment")  # Adjust this list as needed
selected_terms <- c("BC_status1")  # Adjust this list as needed


# Step 2: Extract coefficients, p-values, and confidence intervals from each Cox model
extract_cox_model_info <- function(model, model_name) {
  summary_info <- summary(model)
  coef_table <- coef(summary_info)
  p_value_table <- coef_table[, "Pr(>|z|)"]
  
  # Extract confidence intervals
  conf_interval_table <- confint(model)
  conf_low <- exp(conf_interval_table[, 1])
  conf_high <- exp(conf_interval_table[, 2])
  
  # Combine the extracted information with the model name
  model_info <- data.frame(
    model = model_name,
    predictor = rownames(coef_table),
    coef = coef_table[, 2], ### exp(coef) is the HR
    conf_low = conf_low,
    conf_high = conf_high,
    p_value = p_value_table

  )
  
  # Format the p-values and include significance levels
  formatted_p_values <- as.numeric(format.pval(p_value_table, digits = 10))
  formatted_p_values <- rstatix::p_round(formatted_p_values, digits = 10) ## round p values?
  significance_levels <- ifelse(p_value_table < 0.001, "***",
                                 ifelse(p_value_table < 0.01, "**",
                                        ifelse(p_value_table < 0.05, "*", "")))
  
  # Combine the formatted p-values and significance levels with the extracted information
  # model_info$p_value <- formatted_p_values
  model_info$p_value <- p_value_table
  model_info$significance <- significance_levels
  return(model_info)
}

# Step 3: Extract information from each Cox model and add the model name
combined_cox_info <- data.frame()  # Initialize an empty data frame for the combined results

# Extract information from each named Cox model and add it to the combined table
for (model_name in names(cox_models)) {
  cox_model <- cox_models[[model_name]]
  cox_model_info <- extract_cox_model_info(cox_model, model_name)
  combined_cox_info <- bind_rows(combined_cox_info, cox_model_info)
}

# Filter the combined table to include only the selected terms
filtered_cox_combined_info <- combined_cox_info %>%
  filter(predictor %in% selected_terms)

# Print the filtered combined table with model names, selected predictor names, coefficients, p-values, and confidence intervals
filtered_cox_combined_info$p_value <- as.character(filtered_cox_combined_info$p_value)
filtered_cox_combined_info <- filtered_cox_combined_info %>% mutate_if(is.numeric, ~round(., 2))

View(filtered_cox_combined_info)

write.table(filtered_cox_combined_info, file = "cmr-abstract_LOGISTIC_regressions_table.tsv", sep = "\t", row.names = F)
```



## Linear Regression Models for CMR measures (BC vs no BC)

```{r lin_regression_model_BCvsNoBC}

lin_models <- list()

### max_LV_WT Model
formula = max_LV_WT ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "max_LV_WT-Model"
lin_models <- c(lin_models, tmp)

### LVMi Model
formula = LV_mass2 ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "LVMi-Model"
lin_models <- c(lin_models, tmp)


### LVEDVI MODEL 
formula = LVEDVI ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "LDEVDI-Model"
lin_models <- c(lin_models, tmp)
# summary(model1)


### LVM:LDEDV Model
formula = LVM_LDEDV ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "LVM_LDEDV-Model"
lin_models <- c(lin_models, tmp)


### LVSVi Model
formula = LVSVi ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "LVSVi-Model"
lin_models <- c(lin_models, tmp)


### LVEF Model
formula = LVEF ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "LVEF-Model"
lin_models <- c(lin_models, tmp)

### LVGFI Model
formula = LVGFI ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "LVGFI-Model"
lin_models <- c(lin_models, tmp)

### GLS Model
formula = GLS ~ BC_status + age_at_recruitment + sex +
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "GLS-Model"
lin_models <- c(lin_models, tmp)

### t1 Model
formula = t1 ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "t1-Model"
lin_models <- c(lin_models, tmp)


### max_LAVi Model
formula = max_LAVi ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "max_LAVi-Model"
lin_models <- c(lin_models, tmp)

### LAEF Model
formula = LAEF ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "LAEF-Model"
lin_models <- c(lin_models, tmp)


### RVEDVi Model
formula = RVEDVi ~ BC_status + age_at_recruitment + sex + 
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "RVEDVi-Model"
lin_models <- c(lin_models, tmp)


### RVSVi Model
formula = RVSVi ~ BC_status + age_at_recruitment + sex +
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "RVSVi-Model"
lin_models <- c(lin_models, tmp)


### RVEF Model
formula = RVEF ~ BC_status + age_at_recruitment + sex +
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "RVEF-Model"
lin_models <- c(lin_models, tmp)

### DAo Model
formula = DAo ~ BC_status + age_at_recruitment + sex +
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension
model1 <- glm(formula, data = cmr_df)
tmp <- list(model1)
names(tmp) <- "DAo-Model"
lin_models <- c(lin_models, tmp)
```

### Store lin. regression results in table (BC no BC)
```{r lin_models_storing_BCvsNoBC}

# Step 1: Create a list of Cox models (you should have your own list)
print(length(lin_models))

# Step 2: Extract coefficients and p-values from each model, keeping track of model names
extract_model_info <- function(model, model_name) {
  summary_info <- summary(model)
  coef_table <- coef(summary_info)
  p_value_table <- coef_table[, "Pr(>|t|)"]
  # Extract confidence intervals
  conf_interval_table <- confint(model)
  conf_low <- conf_interval_table[, 1]
  conf_high <- conf_interval_table[, 2]
  
  # Combine the extracted information with the model name
  model_info <- data.frame(
    model = model_name,
    predictor = rownames(coef_table),
    coef = coef_table[, "Estimate"],
    conf_low = conf_low,
    conf_high = conf_high,
    p_value = p_value_table
  )
  
  # Format the p-values and include significance levels
  # formatted_p_values <- as.numeric(format.pval(p_value_table, digits = 3))
  # formatted_p_values <- rstatix::p_round(formatted_p_values, digits = 3) ## round p values?
  significance_levels <- ifelse(p_value_table < 0.001, "***",
                                 ifelse(p_value_table < 0.01, "**",
                                        ifelse(p_value_table < 0.05, "*", "")))
  
  # Combine the formatted p-values and significance levels with the extracted information
  model_info$p_value <- p_value_table
  model_info$significance <- significance_levels
  return(model_info)
}

# Step 3: Extract information from each linear regression model
combined_info <- data.frame()  # Initialize an empty data frame for the combined results

# Extract information from each named model and add it to the combined table
for (model_name in names(lin_models)) {
  model <- lin_models[[model_name]]
  model_info <- extract_model_info(model, model_name)
  combined_info <- bind_rows(combined_info, model_info)
}

# Specify the list of selected terms (predictors) to filter the combined table
# selected_terms <- c("BC_status1", "sex", "age_at_recruitment")  # Adjust this list as needed
# selected_terms <- c("(Intercept)", "BC_status1")  # Adjust this list as needed
selected_terms <- c("BC_status1")  # Adjust this list as needed

# Filter the combined table to include only the selected terms
filtered_combined_info <- combined_info %>%
  filter(predictor %in% selected_terms)

# Print the filtered combined table with model names, selected predictor names, coefficients, p-values, and confidence intervals

filtered_combined_info$p_value <- as.character(filtered_combined_info$p_value)
filtered_combined_info <- filtered_combined_info %>% mutate_if(is.numeric, ~round(., 2))

# Calculate standard errors using p-values
alpha = 0.05  # Significance level

# Calculate z-scores from p-values
filtered_combined_info$z_score <- qnorm(1 - alpha / 2)  # Two-tailed

# Calculate standard errors
filtered_combined_info$std.error <- abs(filtered_combined_info$coef / filtered_combined_info$z_score)
# filtered_combined_info$predictor <- NULL

View(filtered_combined_info)

# write.table(filtered_combined_info, file = "results_lin_regressions_table.tsv", sep = "\t", row.names = F)
# saveRDS(filtered_combined_info, file = "results_lin_regressions_table.rds")
```


```{r}

formula = t1 ~ BC_status + age_at_recruitment + 
  sex +
  smoking_status_at_recruitment + alc_intake_at_recruitment + 
  mets_imaging_visit1 + 
  deprivation_index +
  prev_status_diabetes + prev_status_cholesterol + prev_status_hypertension + 
  LVEF + GLS
model1 <- glm(formula, data = cmr_df)
summary(model1)
```


## Plots for CMR measures regression results

```{r}
### code snippet from LILIANA

# Create a function to make the plots
plot_type <- function(df, chamber_vars, chamber_name) {
  # Filter the dataframe based on the cardiac chamber
  df <- df %>% filter(cmr_var %in% chamber_vars)

  # Convert cmr_var to a factor and specify the levels
  df$cmr_var <- factor(df$cmr_var, levels = rev(chamber_vars))

  # Create the plot
  p <- ggplot(data = df, aes(x = cmr_var, y = estimate, color = term)) +
    geom_point() +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
    scale_color_manual(values = c("#28ae80", "#2c728e", "#472d7b")) +
    labs(title = paste0(chamber_name, " association with sports type"),
         x = "CMR Variable", y = "Beta coefficient", color = "Sport type") +
    theme_minimal() +
    theme(legend.position = "right") +
    coord_flip()

  return(p)
}


LV_vars = c("LVEDVI", "LV_mass2", "GLS")
# Now you can call the function as before
LV_plot <- plot_type(cmr_df, LV_vars, "LV")
# RV_plot <- plot_type(sport_type_results, RV_vars, "RV")
# LA_plot <- plot_type(sport_type_results, LA_vars, "LA")
# RA_plot <- plot_type(sport_type_results, RA_vars, "RA")



# Display the plots
gridExtra::grid.arrange(LV_plot, RV_plot, LA_plot, RA_plot, ncol = 1)
```


```{r}

forest_plot <- ggplot(filtered_combined_info %>% filter(grepl("1", model, ignore.case = TRUE)) , 
                      aes(y = predictor, x = coef)) +
  geom_point() +
  geom_errorbarh(aes(xmin = coef - 1.96 * std.error, xmax = coef + 1.96 * std.error)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
  # coord_flip() +
  facet_wrap(~model, ncol = 1, scales = "free_y") +  # Separate plots for each model
  labs(title = "Forest Plot of Regression Coefficients",
       x = "Coefficient Estimate",
       y = "Variable") +
  theme_minimal()  # Customize the theme as needed

print(forest_plot)

data <- filtered_combined_info %>% filter(grepl("2", model, ignore.case = TRUE))

install.packages("forestplot")

library(forestplot)
options(forestplot-symbolsize = 1.5)  # Adjust the value as needed

forestplot(data$predictor, data$coef, data$conf_low, data$conf_high, 
           pvalues = data$p_value, zero = 0, is.summary=c(TRUE, FALSE, FALSE))
                           
forest_plot <- ggplot(filtered_combined_info %>% filter(grepl("2", model, ignore.case = TRUE)) , 
                      aes(y = predictor, x = coef)) +
  geom_point() +
  geom_errorbarh(aes(xmin = coef - 1.96 * std.error, xmax = coef + 1.96 * std.error)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
  # coord_flip() +
  facet_wrap(~model, ncol = 1, scales = "free_y") +  # Separate plots for each model
  labs(title = "Forest Plot of Regression Coefficients",
       x = "Coefficient Estimate",
       y = "Variable") +
  theme_minimal()  # Customize the theme as needed

print(forest_plot)


# """it would be good to have the CMR metric on the left hand side (where you currently have ("BC_status"). And to have a vertical marker at zero - where we know the effect is null. And also some way to highlight the results that are stat sig"""
```

